---
title: "readTimCottier"
output: html_document
date: "2025-12-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 

Can’t test against a constant in quid .

Easiest thing:

Compare unconstrainted model to the dominance model , which is that each individual has a positive.

But to do that, still can't use their library because it's set up for two conditions.

The way Haaf and Rouder define it, the hypothesis is about the sample not the population. Not important for the direction we are going in.

The standardized effect size of the FLE , the average relative to residuals (measurement error or between-session variance).  In the model will go a point estimate and a prior on the between-session variance distribution, assuming some could be negative (the unconstrained model).



```{r cars}
#install.packages("R.matlab")
#install.packages("quid") 
library(R.matlab)
library(quid) ##Haaf Rouder qualitative difference package
library(tidyverse)
```
## Show how it works with the built-in Stroop example, from the vignette:

The data set contains three columns of interest for our analysis: rtS is the dependent variable, which is the response time in seconds; ID is the participant ID; and cond, a factor with two levels indicating the condition (1 = congruent, 2 = incongruent). This study is a repeated measures design, and all participants completed several trials in both conditions.

```{r stroopy, echo=FALSE}

data(stroop)

resStroop <- constraintBF(formula = rtS ~ ID + cond + ID:cond,
                          data = stroop,
                          whichRandom = "ID",
                          ID = "ID",
                          whichConstraint = c("cond" = "2 > 1"),
                          rscaleEffects = c("ID" = 1, "cond" = 1/6, "ID:cond" = 1/10))

```

The data set contains three columns of interest for our analysis: rtS is the dependent variable, which is the response time in seconds; ID is the participant ID; and cond, a factor with two levels indicating the condition (1 = congruent, 2 = incongruent). This study is a repeated measures design, and all participants completed several trials in both conditions.


## Load our data


```{r readData, echo=FALSE}

FLE<- readMat("FLE_Values(JOV).mat")

overallOfBothSessions<- FLE$FLE.Overall
overallOfBothSessions<- as_tibble(overallOfBothSessions)

sess1JoV<-  as_tibble(FLE$FLE.Sess1)
sess2JoV<-  as_tibble(FLE$FLE.Sess2)

#European Journal of Neuroscience 2025 paper
#Cottier_et_al_2023_Data(EJN).csv

```

Give data better column names

Each has three columns: participant number, Column 2 is pixels column 3 is dva

```{r , echo=FALSE}

sess1JoV <- sess1JoV %>%
  rename(
    ID     = V1,
    pixels = V2,
    dva    = V3
  )

sess2JoV <- sess2JoV %>%
  rename(
    ID     = V1,
    pixels = V2,
    dva    = V3
  )


```

Need to re-arrange the data so that session number is a column (make session long rather than wide)

ID is the participant ID; and cond, a factor with two levels indicating the condition (1 = congruent, 2 = incongruent)

```{r , echo=F}
sess1JoV$session <- 1
sess2JoV$session <- 2

myOverallJoV<- rbind(sess1JoV, sess2JoV)
```



```{r , echo=FALSE}
library(ggplot2)

ggplot(myOverallJoV, aes(x=dva)) + geom_histogram() +
  facet_grid(rows=vars(session)) +
  xlab('FLE size (dva)') +
  ggtitle('JoV2022 FLE') 

# ggplot(overallOfBothSessions, aes(x=)) + geom_histogram() +
#   xlab('FLE size (dva)') +
#   ggtitle('JoV2022 FLE') 
```
## Reproduce scatterplot of s1 and s2

```{r }

wide <- myOverallJoV %>%
  select(ID, dva, session) %>%
  pivot_wider(
    names_from = session,
    values_from = dva,
    names_prefix = "dva_"
  )

ggplot(wide, aes(x = dva_1, y = dva_2)) +
  geom_point() +
  labs(
    x = "Session 1 dva",
    y = "Session 2 dva"
  )

```
