---
title: "Inidividual Differences Investigation for Flash Lag Effect"
date: today
date-format: long

author:
  - name: Frederik Aust
    orcid: 0000-0003-4900-788X
    email: frederik.aust@uni-koeln.de
    affiliations:
      - ref: uoc
  - name: Alex Holcombe
    orcid: 0000-0003-2869-0085
    affiliations:
      - ref: uos

affiliations:
  - id: uoc
    name: University of Cologne
    city: Cologne
    country: Germany
  - id: uos
    name: University of Sydney
    city: Sydney
    state: Australia

theme: lumen
toc: true
highlight-style: github
reference-location: margin

html:
  standalone: true
  embed-resources: true
  self-contained: true
  link-external-icon: true
  citations-hover: true
  footnotes-hover: true
---

```{r}
#| label: setup
#| include: false

library("dplyr")
library("ggplot2")
library("patchwork")
library("papaja")

library("lme4")
library("lmerTest")

library("rstan")
library("tidybayes")

library("logspline")
library("bridgesampling")
```

```{r}
#| label: load-data

fle_data <- read.csv("./publishedMotnPostnDataToLoad/myOverallJoV_readyForQualitativeTest.csv")
```

For this analysis, we need multiple observations per participant to partition the observed variance in degrees of visual angle (`dva`) into true individual differences $\sigma_\theta$ and error variance $\sigma_\epsilon$, i.e. variance between observations within a participant.
Hence, we collapse the data across the two sessions and estimate an overall mean (`(Intercept)`, $\mu_\theta$) and individual differences around this mean for each participant (random effect `ID`, $\sigma_\theta$).

```{r}
#| label: fit-lme-model

fle_lmer <- lmer(dva ~ 1 + (1 | ID), data = fle_data)

summary(fle_lmer)
```

Based on the assumption of a normal distribution of individual differences $\theta_i \sim \mathcal{N}(\mu_\theta, \sigma_\theta)$, we can compute the proportion of participants with truely negative effects $p(\theta_i<0)$ as follows:

$$
p(\theta_i<0) = \Phi\left(\frac{- \mu_\theta}{\sigma_\theta}\right)
$$

```{r}
#| label: lme-proportion-negative

mle_p_negative <- pnorm(
  q = 0
  , mean = fixef(fle_lmer)[1]
  , sd = attr(VarCorr(fle_lmer)$ID, "stddev")
)

mle_p_negative
```

To test whether this proportion is credibly greater than zero, we will fit a Bayesian hierarchical model next.

# Bayesian Hierarchical Model

Following the standard hierarchical model structure for individual differences, we define the following model:

$$
\begin{aligned}
y_{ij} & \sim \mathcal{N}(\theta_i, \sigma_\epsilon) \\
\theta_i & \sim \mathcal{N}(\mu_\theta, \sigma_\theta) \\ \\
\mu_\theta & \sim \log \mathcal{N}(a_\mu, b_\mu) \\ \\
\sigma_\theta & = -\mu_\theta / \Phi^{-1}(0.5 \; p(\theta_i<0)) \\
p(\theta_i<0) & = \Phi(p^{*}_{\theta_i<0}) \\
p^{*}(\theta_i<0) & \sim \mathcal{N}(a_p, b_p) \\ \\
\sigma_\epsilon^2 & \propto 1/\sigma_\epsilon^2
\end{aligned}
$$

We place a log-normal prior on the group mean $\mu_\theta$ to ensure positive values (but we could replace this by something more flexible).
We parameterize the individual differences standard deviation $\sigma_\theta$ in terms of the proportion of individuals that show a negative flash lag effect (i.e., a flash *lead* effect) $p(\theta_i<0)$ to directly estimate this quantity of interest.
To improve sampling efficiency, we place a normal prior on the probit-transformed quantity $p^{*}(\theta_i<0)$.
Finally, we place a non-informative prior on the error variance $\sigma_\epsilon^2$.


## Prior distributions

```{r}
#| label: plot-prior-pse

p_negative_prior <- rnorm(1e5, mean = -2/3, sd = 2/3) |>
  pnorm(0) |>
  (\(x) x / 2)()

p_negative_prior |>
  logspline::logspline(lbound = 0, ubound = 0.5) |>
  (
    \(fit) {
      curve(
        logspline::dlogspline(x, fit)
        , from = 0, to = 0.5
        , lwd = 2
        , xlab = expression(p[theta[i] < 0])
        , ylab = "Density"
        , las = 1
        , n = 1001
      )
    }
  )()
  

curve(
  dlnorm(x, meanlog = 0.5, sdlog = 1)
  , from = 0, to = 5
  , lwd = 2
  , xlab = expression(mu[theta])
  , ylab = "Density"
  , las = 1
  , n = 1001
)

mu_prior <- rlnorm(1e5, meanlog = 0.5, sdlog = 1)
sigma_prior <- -mu_prior / qnorm(p_negative_prior)

sigma_prior |>
  logspline::logspline(lbound = 0) |>
  (
    \(fit) {
      curve(
        logspline::dlogspline(x, fit)
        , from = 0, to = 5
        , lwd = 2
        , xlab = expression(sigma[theta])
        , ylab = "Density"
        , las = 1
        , n = 1001
      )
    }
  )()

unconstrained_joint_prior <- data.frame(mu = mu_prior, sigma = sigma_prior, p_negative = p_negative_prior) |>
  ggplot() +
    aes(x = mu, y = sigma, color = p_negative) +
    geom_point(alpha = 0.1) +
    stat_density_2d(color = "white", bins = 10) +
    geom_abline(intercept = 0, slope = 1, color = "white", linetype = "22") +
    # annotate("point", x = fixef(fle_lmer)[1], y = attr(VarCorr(fle_lmer)$ID, "stddev"), shape = 21, color = "black", fill = "white", size = 5) +
    labs(
      x = expression(mu[theta])
      , y = expression(sigma[theta])
      , color = expression(p(theta[i] < 0))
    ) +
    scale_color_viridis_c(option = "C", limits = c(0, 0.5)) +
    lims(x = c(0, 5), y = c(0, 5)) +
    papaja::theme_apa(base_size = 16)

unconstrained_joint_prior
```


## Model estimation

```{r}
#| label: fit-bayesian-model

pse_model <- readLines("./models/pse-unconstrained.stan") |>
  paste(collapse = "\n")

fle_stan_data <- list(
  N = nrow(fle_data) |> as.integer()
  , n_subjects = length(unique(fle_data$ID)) |> as.integer()
  , id = as.integer(as.factor(fle_data$ID))
  , y = fle_data$dva

  # Prior settings
  , pi_mean = -2/3
  , pi_sd = 2/3
  , mu_theta_mean = 0.5
  , mu_theta_sd = 1
)

fle_bayes <- rstan::stan(
  model_code = pse_model
  , data = fle_stan_data
  , iter = 1e5/4 + 1000
  , warmup = 1000
  , chains = 4
  , cores = 4
  , seed = 123
)

# model diagnostics
str_post_unconstrained <- data.frame(summary(fle_bayes)$summary)
check_hmc_diagnostics(fle_bayes)

```

@tbl-compare-estimates shows the close agreement between the posterior means of the parameter estimates to the point estimates from the (restricted) maximum likelihood estimation via `lmer()`.


```{r}
#| label: tbl-compare-estimates
#| tbl-cap: "Comparison of Bayesian posterior means and MLE estimates."

data.frame(
  parameter = c("mu_theta", "sigma_theta", "sigma_epsilon", "p_negative_theta")
  , bayes = fle_bayes |>
    summary(pars = c("mu_theta", "sigma_theta", "sigma_epsilon", "p_negative_theta")) |>
    (\(x) x$summary)() |>
    as.data.frame() |>
    dplyr::pull("mean")
  , mle = c(
    fixef(fle_lmer)[1],
    attr(VarCorr(fle_lmer)$ID, "stddev"),
    sigma(fle_lmer),
    mle_p_negative
  )
) |>
  knitr::kable()
```

```{r}
#| label: bayesian-model-summary

fle_bayes |>
  summary(pars = c("mu_theta", "sigma_theta", "sigma_epsilon", "p_negative_theta")) |>
  (\(x) x$summary)() |>
  tidyr::as_tibble(rownames = "parameter") |>
  dplyr::mutate(
    parameter = gsub("mu", "$\\\\mu", parameter) |>
      gsub("sigma", "$\\\\sigma", x = _) |>
      gsub("_(.+)", "_{\\\\\\1}$", x = _) |>
      gsub("^p.+", "$p(\\\\theta_i<0)$", x = _)
  ) |>
  knitr::kable(digits = 3)
```

```{r}
#| label: fig-posterior-distributions
#| fig-cap: "Posterior distributions for the group mean of individual differences $\\mu_{\\theta}$, the standard deviation of individual differences $\\sigma_{\\theta}$, and the proportion of participants with a negative flash lag effect $p(\\theta_i<0)$."

(tidybayes::gather_draws(
  fle_bayes
  , mu_theta, sigma_theta, p_negative_theta
) |>
  dplyr::mutate(
    .variable = factor(
      .variable
      , levels = c("mu_theta", "sigma_theta", "p_negative_theta")
      , labels = c("mu[theta]", "sigma[theta]", "p(theta[i] < 0)")
    )
  ) |>
  ggplot() +
  aes(x = .value, fill = after_stat(level)) +
  stat_halfeye(fill_type = "segments", normalize = "panels", .width = c(.5, .8, .95, .99, 1), slab_color = "black", alpha = 0.7, fatten_point = 2, interval_size_range = c(1, 2)) +
  annotate("point", x = c(0, 0.5), y = 0, color = "white") + 
  scale_fill_viridis_d(option = "B", begin = 0.9, end = 0.1) +
  guides(
    fill = guide_legend(
      title = "Credible level"
      , override.aes = list(color = NA)
      , reverse = TRUE
      , direction = "horizontal"
      , title.position = "top"
      , title.hjust = 0.5
    )
  ) +
  coord_cartesian(ylim = c(-0.05, NA)) +
  facet_wrap(
    ~ .variable
    , scales = "free"
    , labeller = label_parsed
    , ncol = 2
  ) +
  labs(x = "Parameter value") +
  papaja::theme_apa(base_size = 16) +
  theme(
    axis.line.y = element_blank()
    , axis.ticks.y = element_blank()
    , axis.text.y = element_blank()
    , axis.title.y = element_blank()
    # , legend.title.position = "top"
  )) |>
  lemon::reposition_legend("center", panel = 'panel-2-2')
```

### Estimates of individual effects

```{r}
#| label: fig-individual-effects

theta_i <- tidybayes::gather_draws(
  fle_bayes
  , theta[i]
)

# i as factor sorted by median of posterior distribution
theta_i_ordered <- tidybayes::median_qi(theta_i) |>
  mutate(
    i = factor(
      i
      , labels = unique(i)[order(.value)]
      , levels = unique(i)[order(.value)]
    )
  )

theta_caterpillar <- theta_i |>
  dplyr::mutate(
    # i as factor sorted by median of posterior distribution
    ID = factor(
      i
      , labels = unique(theta_i_ordered$i)[order(theta_i_ordered$.value)]
      , levels = unique(theta_i_ordered$i)[order(theta_i_ordered$.value)]
    )
  ) |>
  dplyr::rename(theta = .value) |>
  ggplot() +
    aes(x = ID, y = theta) +
    geom_hline(yintercept = 0, color = "black", linetype = "22") +
    stat_pointinterval(aes(fill = after_stat(y) < 0, interval_color = after_stat(y) < 0), point_color = "white", .width = c(.95), fatten_point = 2, linewidth = 1, shape = 21) +
    scale_fill_viridis_d(option = "B", begin = 0.9, end = 0.1, aesthetics = c("fill", "interval_color")) +
    guides(interval_color = "none", fill = "none", linewidth = "none") +
    labs(y = expression(theta), x = "Participant ID") +
    coord_cartesian(ylim = c(-5, 7.5)) +
    papaja::theme_apa(base_size = 16) +
    theme(
      # axis.line.x = element_blank()
      # , axis.ticks.x = element_blank()
      , axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1)
    )
```

```{r}
#| label: fig-posterior-population-distribution

posterior_population_distribution <- tidybayes::gather_draws(
  fle_bayes
  , mu_theta, sigma_theta
) |>
  tidyr::nest(data = c(.chain, .iteration, .draw)) |>
  tidyr::pivot_wider(names_from = .variable, values_from = .value) |>
  dplyr::mutate(
    density = purrr::map2(
      mu_theta
      , sigma_theta
      , \(mu, sigma) {
        dnorm(
          x = seq(-5, 7.5, length.out = 101)
          , mean = mu
          , sd = sigma
        )
      }
    )
    , theta = list(seq(-5, 7.5, length.out = 101))
  )

theta_dist <- posterior_population_distribution |>
  tidyr::unnest(cols = c(density, theta)) |>
  ggplot() +
    aes(x = theta, y = density) +
    geom_vline(xintercept = 0, color = "black", linetype = "22") +
    geom_hline(yintercept = 0, color = "black") +
    geom_lineribbon(
      aes(ymin = after_stat(ymin), ymax = after_stat(ymax))
      , stat = "summary"
      , fun.data = "mean_qi"
      , color = "black"
      , fill = "black"
      , alpha = 0.2
    ) +
    # geom_rug(
    #   data = tidybayes::median_qi(theta_i)
    #   , aes(x = .value, y = 0)
    #   , sides = "b"
    #   , color = "black"
    #   , alpha = 0.5
    # ) +
    # geom_point(
    #   data = tidybayes::median_qi(theta_i)
    #   , aes(x = .value, y = 0, fill = .value < 0)
    #   , color = "white"
    #   , shape = 21
    #   , size = 2
    # ) +
    scale_fill_viridis_d(option = "B", begin = 0.9, end = 0.1) +
    guides(fill = "none") +
    labs(x = expression(theta), y = "Density") +
    papaja::theme_apa(base_size = 16)
```

```{r}
#| label: fig-posterior-theta-distribution
#| fig-cap: "Posterior distribution of individual effects $\\theta_i$ for each participant (left) and the posterior distribution of the population distribution of individual effects $\\theta_i$ (right). The horizontal dashed line marks the point of no effect, i.e., $\\theta_i = 0$."

theta_caterpillar + (theta_dist + coord_flip() + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())) +
  plot_layout(ncol = 2, widths = c(1, 1/3), guides = "collect", axes = "collect_y")
```

## Bayes factor approximations

Finally, we compute the Bayes factor in favor of the assumption of an unconstrained population distribution ($p(\theta_i<0) > 0$) over the strictly positive population distribution (stochastic dominance).
One approach is to approximate this Bayes factor using the Savage-Dickey density ratio for $p(\theta_i<0) = 0$.

```{r}
#| label: fig-bayes-factor
#| fig-cap: "Posterior and prior density for the proportion of participants with a negative flash lag effect, $p(\\theta_i<0)$. The Bayes factor in favor of the alternative hypothesis ($p(\\theta_i<0) > 0$) over the null hypothesis ($p(\\theta_i<0) = 0$) is given by the ratio of the prior density to the posterior density at $p(\\theta_i<0) = 0$."

# Extract posterior samples for p_negative_theta
p_negative_draws <- tidybayes::gather_draws(fle_bayes, p_negative_theta)

# logspline approximation of posterior density
p_negative_draws_ls <- logspline::logspline(
  p_negative_draws$.value
  , lbound = 0, ubound = 0.5
)

# logspline approximation of prior density
prior_ls <- rnorm(1e5, mean = fle_stan_data$pi_mean, sd = fle_stan_data$pi_sd) |>
  pnorm() |>
  (\(x) x / 2)() |>
  logspline::logspline(lbound = 0, ubound = 0.5)

prior_density <- logspline::dlogspline(0, prior_ls)
posterior_density <- logspline::dlogspline(0, p_negative_draws_ls)


curve(
  logspline::dlogspline(x, p_negative_draws_ls)
  , from = 0, to = 0.5
  , col = "black", lwd = 3
  , main = "Posterior Density of P(p_negative_theta)"
  , xlab = "p_negative_theta"
  , n = 1001
)

curve(
  logspline::dlogspline(x, prior_ls)
  , from = 0, to = 0.5
  , col = "grey", lwd = 3
  , add = TRUE
  , n = 1001
)

points(
  x = c(0, 0)
  , y = c(prior_density, posterior_density)
  , col = c("white", "white")
  , bg = c("grey", "black")
  , pch = 21
  , cex = 2
)
```

```{r}
#| label: bayes-factor-calculation

bayes_factor_10 <- prior_density / posterior_density
bayes_factor_10
```

We find overwhelming evidence for the unconstrained population distribution.
In other words, the assumption that noone exhibits a flash *lead* effect, seems highly implausible.

This approach is, however, approximate and conceptually problematic.
Under the assumption of a normal population distribution, the probability of negative effects cannot be 0, $p(\theta_i<0) \neq 0$.
So let's try another approach.

For an exact model comparison, we need to replace the assumed normal population distribution by a distribution family with support over the positive domain (e.g., truncated normal, log-normal, gamma, or inverse gamma distribution).
This raises questions about the most appropriate distribution that lead us down the path of fitting and averaging multiple models.
I will resist that temptation and proceed with a truncated normal distribution for now. 

# Stochastic dominance analysis (?)

Following @doi:10.3758/s13423-018-1522-x : 
Bayesian model comparison with various Bayesian hierarchical models, each defined by different constraints on parameters. 
Unconstrained model without any equality or order constraints on the individual effects is defined above. 

**Positive effect model (dominance model)**: 
Order constraint, constrains individual effects to be positive, but effects differ/vary interindividually. Truncated normal distribution as suggested by literature.

Defining $\sigma_{\theta}$ the same way as in the unconstrained model would not make any sense conceptually, since in this model $p(\theta_i<0)$ is assumed to be zero. Therefore chose log-normal as informative prior (as prior of unconstrained model was similar to prior for $\mu_\theta$).

$$
\begin{aligned}
y_{ij} & \sim \mathcal{N}(\theta_i, \sigma_\epsilon) \\ \\
\theta_i & \sim \mathcal{N}^+(\mu_\theta, \sigma_\theta) \\ 
\sigma_\epsilon^2 & \propto 1/\sigma_\epsilon^2 \\ \\
\mu_\theta & \sim \log \mathcal{N}(a_\mu, b_\mu) \\ 
\sigma_\theta & \sim \log \mathcal{N}(a_\sigma, b_\sigma) \\
\end{aligned}
$$

**Common-effect model**: 
No individual variablity in effects, all inidividual share a common, constant effect different from zero (positive).
Unlikely to be an accurate representation of effects in psychology, but used as (1) benchmark to check for individual differences (model could be preferred should individual difference be very small) and (2) design check (design might not be able to capture individual differences).

Priors are only needed for $\theta$ (as there is uncertainty regarding the size of the common effect) and $\sigma_\epsilon^2$. Logarithmic normal distribution as prior for $\theta$, constraining it to positive real numbers (paralleling prior of $\mu_{\theta}$ in unconstrained model). 

$$
\begin{aligned}
y_{ij} & \sim \mathcal{N}(\theta, \sigma_\epsilon) \\ \\
\theta & \sim \log \mathcal{N}(a_\mu, b_\mu) \\
\sigma_\epsilon^2 & \propto 1/\sigma_\epsilon^2
\end{aligned}
$$

**Average null model**:
Average effect across individuals is constrained to be zero, but effects are assumed to vary across individuals. Conceptualized as a more forgiving alternative to the null model (see below), less reactive to contaminated data. $\theta_i$ are modeled as normally distributed around a mean of zero. Only slight deviations from zero align with the idea of this model as an alternative to the null model, therefore steep log-normal distribution with relatively small mean to be chosen as prior for $\sigma_{\theta}$ (see average null vs unconstrained).


$$
\begin{aligned}
y_{ij} & \sim \mathcal{N}(\theta_i, \sigma_\epsilon) \\ \\
\theta_i & \sim \mathcal{N}(0, \sigma_\theta) \\ 
\sigma_\epsilon^2 & \propto 1/\sigma_\epsilon^2 \\ \\
\sigma_\theta & \sim \log \mathcal{N}(a_\sigma, b_\sigma)
\end{aligned}
$$

**Null model**: 
Specifies a true effect of zero for all individuals, most constrained model. Used as benchmark for no effect and no interindividual variability. 
There is no uncertainty regarding the size of the effect, as it is assumed to be zero. So no prior for theta. Only prior for sigma (measurement error around true $\theta = 0$) needed. 

$$
\begin{aligned}
y_{ij} & \sim \mathcal{N}(0, \sigma_\epsilon) \\ \\
\sigma_\epsilon^2 & \propto 1/\sigma_\epsilon^2
\end{aligned}
$$

## (1) Dominance (truncated normal) vs unconstrained

Priors chosen to be similar to priors of unconstrained model, therefore approximately same percentage of theoretically negative $\theta_i$. But this time, negative $\theta_i$ are excluded / impossible by stan model definition (truncated normal distribution).

```{r}
#| label: prior-plot-mu-dominance
#| fig-cap: "Prior distribution of $\\mu_{\\theta}$ in dominance model"

mu_prior_truncn <- rlnorm(1e5, meanlog = 0.5, sdlog = 1)

curve(
  dlnorm(x, meanlog = 0.5, sdlog = 1)
  , from = 0, to = 5
  , lwd = 2
  , xlab = expression(mu[theta])
  , ylab = "Density"
  , las = 1
  , n = 1001
)
```

```{r}
#| label: prior-plot-sigma-dominance
#| fig-cap: "Prior distribution of $\\sigma_{\\theta}$ in dominance model"

sigma_prior_truncn <- rlnorm(1e5, meanlog = 0.5, sdlog = 1)

curve(
  dlnorm(x, meanlog = 0.5, sdlog = 1)
  , from = 0, to = 5
  , lwd = 2
  , xlab = expression(sigma[theta])
  , ylab = "Density"
  , las = 1
  , n = 1001
)
```

```{r}
#| label: fit-dominance-model

pse_dominance_model <- readLines("./models/pse-dominance.stan") |>
  paste(collapse = "\n")

fle_stan_dominance_data <- list(
  N = nrow(fle_data) |> as.integer()
  , n_subjects = length(unique(fle_data$ID)) |> as.integer()
  , id = as.integer(as.factor(fle_data$ID))
  , y = fle_data$dva

  # Prior settings
  , mu_theta_mean = 0.5
  , mu_theta_sd = 1
  , sigma_theta_mean = 0.5
  , sigma_theta_sd = 1
)

fle_dominance_bayes <- rstan::stan(
  model_code = pse_dominance_model
  , data = fle_stan_dominance_data
  , iter = 1e5/4 + 1000
  , warmup = 1000
  , chains = 4
  , cores = 4
  , seed = 123
)

# model diagnostics
str_post_dominance<- data.frame(summary(fle_dominance_bayes)$summary)  
check_hmc_diagnostics(fle_dominance_bayes)

```

### Estimates of individual effects

```{r}
#| label: fig-indiv-effects-truncated

theta_i_truncn <- tidybayes::gather_draws(
  fle_dominance_bayes
  , theta[i]
)

theta_i_trunc_ordered <- tidybayes::median_qi(theta_i_truncn) |>
  mutate(i = factor(i
                    ,levels = unique(i)[order(.value)]        #levels are ordered after median, 64 = min value = first level
                    ,labels = unique(i)[order(.value)]        #I create this solely for creating levels in draws, as geom-f calls on draws
                    ) 
         )

# Plot median estimates
truncated_caterpillar <- theta_i_truncn |>
  mutate(ID = factor(i
                     ,levels = unique(theta_i_trunc_ordered$i)[order(theta_i_trunc_ordered$.value)]  #new var ID with ordered med levels
                     ,labels = unique(theta_i_trunc_ordered$i)[order(theta_i_trunc_ordered$.value)]
                     )
  ) |>
  rename(theta = .value) |>
  ggplot() +
  aes(x = ID, y = theta) +
  geom_hline(yintercept = 0, color = "black", linetype = "22") + #string-linetype: line for 2 units, followed by 2 off
  stat_pointinterval(           #draws point est + CI for draw-input, point_interval function is median_qi by default
    fill = viridisLite::inferno(1, begin = 0.9), interval_color = viridisLite::inferno(1, begin = 0.9), point_color = "white"
    , .width = c(.95), fatten_point = 2, linewidth = 1, shape = 21
  ) +
  labs(y = expression(theta[i]), x = "Participant ID") +
  coord_cartesian(ylim = c(0,6)) +
  papaja::theme_apa(base_size = 16) + 
  theme(
    axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 0)
  )
  
```

```{r}
#| label: fig-post-population-distribution-truncated

post_populDist_truncated <- tidybayes::gather_draws(
  fle_dominance_bayes
  , mu_theta, sigma_theta
) |>
  tidyr::nest(data = c(.chain, .iteration, .draw)) |>
  tidyr::pivot_wider(names_from = .variable, values_from = .value) |>
  mutate(
    density = purrr::map2(                      # iteration of each density function for each pair of mu and sigma
      mu_theta, sigma_theta, 
      \(mu, sigma){
        dnorm(
          x = seq(0, 6, length.out = 101)
          , mean = mu
          , sd = sigma
        )
      }
    )
    ,theta = list(seq(0, 6, length.out = 101))    # corresponsing values on x for plot
  )
  
 theta_trunc_dist <- post_populDist_truncated |>
  tidyr::unnest(cols = c(density, theta)) |>
  ggplot() + 
  aes(x = theta, y = density) +
   geom_vline(xintercept = 0, color = "black", linetype = "22") +
   geom_hline(yintercept = 0, color = "black") + 
   geom_lineribbon(                              # function designed to use with output from point_interval
     aes(ymin = after_stat(ymin), ymax = after_stat(ymax))
     , stat = "summary"
     , fun.data = "median_qi"
     , color = "black"
     , fill = "black"
     , alpha = 0.2
   ) + 
   labs(x = expression(theta[i]), y = "Density") + 
   papaja::theme_apa(base_size = 16)

```

```{r}
#| label: fig-posterior-theta-distribution-truncated
#| fig-cap: "Posterior distribution of individual effects $\\theta_i$ for each participant (left) and the posterior distribution of the population distribution of individual effects $\\theta_i$ (right). The horizontal dashed line marks the point of no effect, i.e., $\\theta_i = 0$."

truncated_caterpillar + (theta_trunc_dist + coord_flip() + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())) +
  plot_layout(ncol = 2, width = c(1, 1/3), axes = "collect_y")
```

### Bayes factor with unconstrained model

```{r}
#| label: bayes-factor-dominance

unconstrained_ml <- bridge_sampler(fle_bayes, method = "warp3", repetitions = 30, silent = TRUE)
dominance_ml <- bridge_sampler(fle_dominance_bayes, method = "warp3", repetitions = 30, silent = TRUE)

# bayes factor
bayes_factor(unconstrained_ml, dominance_ml)

```

Comparing to a strictly positive population distribution (truncated normal in this case), we find overwhelming evidence for an unconstrained population distribution, implying that some individuals could truly exhibit a flash lead effect. In other words, these results suggest that it is highly unlikely that no one exhibits a true lead effect. 


## (2) Common vs unconstrained

```{r}
#| label: prior-plot-common
#| fig-cap: "Prior distribution of $\\theta$ in common model"

mu_prior_common <- rlnorm(1e5, meanlog = 0.5, sdlog = 1)

curve(
  dlnorm(x, meanlog = 0.5, sdlog = 1)
  , from = 0, to = 5
  , lwd = 2
  , xlab = expression(theta)
  , ylab = "Density"
  , las = 1
  , n = 1001
)

```

```{r}
#| label: fit-common-model

pse_common_model <- readLines("./models/pse-common.stan") |>
  paste(collapse = "\n")

fle_stan_common_data <- list(
  N = nrow(fle_data) |> as.integer()
  , n_subjects = length(unique(fle_data$ID)) |> as.integer()
  , id = as.integer(as.factor(fle_data$ID))
  , y = fle_data$dva

  # Prior settings
  , theta_mean = 0.5
  , theta_sd = 1
)

fle_common_bayes <- rstan::stan(
  model_code = pse_common_model
  , data = fle_stan_common_data
  , iter = 1e5/4 + 1000
  , warmup = 1000
  , chains = 4
  , cores = 4
  , seed = 123
)

# model diagnostics
str_post_common <- data.frame(summary(fle_common_bayes)$summary)
check_hmc_diagnostics(fle_common_bayes)
```

```{r}
#| label: fig-posterior-distributions-common
#| fig-cap: "Posterior distributions for the common effect $\\theta$."

tidybayes::gather_draws(
  fle_common_bayes
  , theta
) |>
  ggplot() +
  aes(x = .value, fill = after_stat(level)) +
  stat_halfeye(fill_type = "segments", normalize = "panels", .width = c(.5, .8, .95, .99, 1), slab_color = "black", alpha = 0.7, fatten_point = 2, interval_size_range = c(1, 2)) +
  annotate("point", x = c(0, 0.5), y = 0, color = "white") + 
  scale_fill_viridis_d(option = "B", begin = 0.9, end = 0.1) +
  guides(
    fill = guide_legend(
      title = "Credible level"
      , override.aes = list(color = NA)
      , reverse = TRUE
      , direction = "horizontal"
      , title.position = "top"
      , title.hjust = 0.5
    )
  ) +
  labs(x = expression(theta), y = "Density") +
  papaja::theme_apa(base_size = 16)

```

### Bayes factor with unconstrained model

```{r}
#| label: bayes-factor-common

common_ml <- bridge_sampler(fle_common_bayes, method = "warp3", repetitions = 30, silent = TRUE)

bayes_factor(unconstrained_ml, common_ml)

```

We find overwhelming evidence in favor of the unconstrained model over the common effect model. It therefore seems highly implausible that all participants share a common true effect of the same size, as models allowing for interindividual differences in true effects are a better fit for the data. These results also validate / support the study design, as interindividual differences in true effects were measureable. 

## (3) Average null vs unconstrained

```{r}
#| label: prior-plot-average
#| fig-cap: "Prior distribution of $\\sigma_{\\theta}$ in average null model"

sigma_prior_average <- rlnorm(1e5, meanlog = -0.5, sdlog = 0.5)

curve(
  dlnorm(x, meanlog = -0.5, sdlog = 0.5)
  , from = 0, to = 5
  , lwd = 2
  , xlab = expression(sigma[theta])
  , ylab = "Density"
  , las = 1
  , n = 1001
)

# Mean of prior sigma_theta
exp(-0.5 + (0.5)^2/2)

```

```{r}
#| label: fit-average-model

pse_average_model <- readLines("./models/pse-average-null.stan") |>
  paste(collapse = "\n")

fle_stan_average_data <- list(
  N = nrow(fle_data) |> as.integer()
  , n_subjects = length(unique(fle_data$ID)) |> as.integer()
  , id = as.integer(as.factor(fle_data$ID))
  , y = fle_data$dva

  # Prior settings
  , sigma_theta_mean = -0.5
  , sigma_theta_sd = 0.5
)

fle_average_bayes <- rstan::stan(
  model_code = pse_average_model
  , data = fle_stan_average_data
  , iter = 1e5/4 + 1000
  , warmup = 1000
  , chains = 4
  , cores = 4
  , seed = 123
)

# model diagnostics
str_post_average <- data.frame(summary(fle_average_bayes)$summary)
check_hmc_diagnostics(fle_average_bayes)
```

### Estimates of individual effects

```{r}
#| label: fig-indiv-effects-average

theta_i_average <- tidybayes::gather_draws(
  fle_average_bayes
  , theta[i]
)

theta_i_average_ordered <- tidybayes::median_qi(theta_i_average) |>
  mutate(i = factor(i
                    ,levels = unique(i)[order(.value)]        
                    ,labels = unique(i)[order(.value)]        #I create this solely for creating levels in draws, as geom-f calls on draws
                    ) 
         )

# Plot median estimates
average_caterpillar <- theta_i_average |>
  mutate(ID = factor(i
                     ,levels = unique(theta_i_average_ordered$i)[order(theta_i_average_ordered$.value)]  
                     ,labels = unique(theta_i_average_ordered$i)[order(theta_i_average_ordered$.value)]
                     )
  ) |>
  rename(theta = .value) |>
  ggplot() +
  aes(x = ID, y = theta) +
  geom_hline(yintercept = 0, color = "black", linetype = "22") + #string-linetype: line for 2 units, followed by 2 off
  stat_pointinterval(           #draws point est + CI for draw-input, point_interval function is median_qi by default
    aes(fill = after_stat(y) < 0, interval_color = after_stat(y) < 0), 
    point_color = "white", .width = c(.95), fatten_point = 2, linewidth = 1, shape = 21
  ) +
  scale_fill_viridis_d(option = "B", begin = 0.9, end = 0.1, aesthetics = c("fill", "interval_color")) +
  guides(interval_color = "none", fill = "none", linetype = "none") +
  labs(y = expression(theta[i]), x = "Participant ID") +
  coord_cartesian(ylim = c(-4,7)) +
  papaja::theme_apa(base_size = 16) + 
  theme(
    axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 0)
  )
  
```

```{r}
#| label: fig-post-population-distribution-average

post_populDist_average <- tidybayes::gather_draws(
  fle_average_bayes
  , sigma_theta
) |>
  tidyr::nest(data = c(.chain, .iteration, .draw)) |>
  tidyr::pivot_wider(names_from = .variable, values_from = .value) |>
  mutate(
    mu_theta = 0        
  ) |>
  mutate(
    density = purrr::map2(                      # iteration of each density function for each pair of mu and sigma
      mu_theta, sigma_theta, 
      \(mu, sigma){
        dnorm(
          x = seq(-4, 7, length.out = 101)
          , mean = mu
          , sd = sigma
        )
      }
    )
    ,theta = list(seq(-4, 7, length.out = 101))    # corresponsing values on x for plot
  )
  
 theta_average_dist <- post_populDist_average |>
  tidyr::unnest(cols = c(density, theta)) |>
  ggplot() + 
  aes(x = theta, y = density) +
   geom_vline(xintercept = 0, color = "black", linetype = "22") +
   geom_hline(yintercept = 0, color = "black") + 
   geom_lineribbon(                              # function designed to use with output from point_interval
     aes(ymin = after_stat(ymin), ymax = after_stat(ymax))
     , stat = "summary"
     , fun.data = "median_qi"
     , color = "black"
     , fill = "black"
     , alpha = 0.2
   ) + 
   scale_fill_viridis_d(option = "B", begin = 0.9, end = 0.1) +
   guides(fill = "none") +
   labs(x = expression(theta[i]), y = "Density") + 
   papaja::theme_apa(base_size = 16)

```

```{r}
#| label: fig-posterior-theta-distribution-ave rage
#| fig-cap: "Posterior distribution of individual effects $\\theta_i$ for each participant (left) and the posterior distribution of the population distribution of individual effects $\\theta_i$ (right). The horizontal dashed line marks the point of no effect, i.e., $\\theta_i = 0$."

average_caterpillar + (theta_average_dist + coord_flip() + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())) +
  plot_layout(ncol = 2, width = c(1, 1/3), axes = "collect_y", guides = "collect")
```

Frage: 
Unsicher ob density so sinnvoll oder richtig? 

Vllt bin ich damit komplett auf dem Holzweg, aber kann ich post-mu und post-sigma aus den theta berechnen? Dachte zuerst vllt mu = mean der true-scores pro Iteration (weil theta = true score ja eig Messfehler-bereinigt) aber das passt bei Modellen mit mu-Schätzungen nicht. 

Auch Implikationen des oben gezeichneten density plots für mich schwer zu beurteilen, weil ich mit Doku nicht 100% nachempfinden konnte, wie lineribbon die CI berechnet (und deren Verlauf ist iwie strange). Einfach nochmal stärkerer Hinweis darauf, dass Annahme symmetrischer Populationsverteilung um 0 laut Daten fehlgeleitet? Oder Sprezifikations-/ konzeptueller Fehler?

### Bayes factor with unconstrained model

```{r}
#| label: bayes-factor-average

average_ml <- bridge_sampler(fle_average_bayes, method = "warp3", repetitions = 30, silent = TRUE)

bayes_factor(unconstrained_ml, average_ml)

```

## (4) Null vs unconstrained

```{r}
#| label: fit-null-model

pse_null_model <- readLines("./models/pse-null.stan") |>
  paste(collapse = "\n")

fle_stan_null_data <- list(
  N = nrow(fle_data) |> as.integer()
  , n_subjects = length(unique(fle_data$ID)) |> as.integer()
  , id = as.integer(as.factor(fle_data$ID))
  , y = fle_data$dva
)

fle_null_bayes <- rstan::stan(
  model_code = pse_null_model
  , data = fle_stan_null_data
  , iter = 1e5/4 + 1000
  , warmup = 1000
  , chains = 4
  , cores = 4
  , seed = 123
)

# model diagnostics
str_post_null <- data.frame(summary(fle_null_bayes)$summary)
check_hmc_diagnostics(fle_null_bayes)
```

### Bayes factor with unconstrained model

```{r}
#| label: bayes-factor-null

null_ml <- bridge_sampler(fle_null_bayes, method = "warp3", repetitions = 30, silent = TRUE)

bayes_factor(unconstrained_ml, null_ml)
```

# Alternative dominance models
## Log-normal distributed $\theta_i$

We have now been tempted to return to the question of the most appropriate distribution over the positive domain. An alternative to the previously fitted truncated normal distribution is the log-normal distribution. We define the following model:

$$
\begin{aligned}
y_{ij} & \sim \mathcal{N}(\theta_i, \sigma_\epsilon) \\ \\
\theta_i & \sim \log \mathcal{N}(\mu_\theta, \sigma_\theta) \\
\sigma_\epsilon^2 & \propto 1/\sigma_\epsilon^2 \\ \\
\mu_\theta & = \ln \left( \frac{\bar{x}}{\sqrt{s^2+\bar{x}^2}} \right) \\
\sigma_\theta & = \ln \left( \frac{s^2}{\bar{x}^2} + 1  \right) \\ \\
\bar{x} & \sim \log \mathcal{N}(a_x, b_x) \\
s & \sim \log \mathcal{N}(a_s, b_s)
\end{aligned}
$$

In this model priors are put on the mean ($\bar{x}$) and the standard deviation ($s$) of the log-normal distribution. We chose this approach because we assume that expressing prior beliefs about the mean and standard deviation of $\theta_i$ is more intuitive than expressing prior beliefs about the parameters $\mu$ and $\sigma$ of the log-normal distribution.  

```{r}
#| label: fit-logN-model

mean_prior_log <- rlnorm(1e5, meanlog = 0.5, sdlog = 1)
sd_prior_log <-rlnorm(1e5, meanlog = 0.5, sdlog = 1)

pse_logN_model <- readLines("./models/pse-dominance-logN.stan") |>
  paste(collapse = "\n")

fle_stan_logN_data <- list(
  N = nrow(fle_data) |> as.integer()
  , n_subjects = length(unique(fle_data$ID)) |> as.integer()
  , id = as.integer(as.factor(fle_data$ID))
  , y = fle_data$dva
  
  # Prior settings
  , mean_logN_mu = 0.5
  , mean_logN_sigma = 1
  , sd_logN_mu = 0.5
  , sd_logN_sigma = 1
)

fle_logN_bayes <- rstan::stan(
  model_code = pse_logN_model
  , data = fle_stan_logN_data
  , iter = 1e5/4 + 1000
  , warmup = 1000
  , chains = 4
  , cores = 4
  , seed = 123
)

# model diagnostics
str_post_logN <- data.frame(summary(fle_logN_bayes)$summary)
check_hmc_diagnostics(fle_logN_bayes)
```

### Estimated of individual effects 

```{r}
#| label: fig-indiv-effects-log-normal

theta_i_logN <- tidybayes::gather_draws(
  fle_logN_bayes
  , theta[i]
)

theta_i_logN_ordered <- tidybayes::median_qi(theta_i_logN) |>
  mutate(i = factor(i
                    ,levels = unique(i)[order(.value)]        
                    ,labels = unique(i)[order(.value)]        #I create this solely for creating levels in draws, as geom-f calls on draws
                    ) 
         )

# Plot median estimates
log_normal_caterpillar <- theta_i_logN |>
  mutate(ID = factor(i
                     ,levels = unique(theta_i_logN_ordered$i)[order(theta_i_logN_ordered$.value)]  #new var ID with ordered med levels
                     ,labels = unique(theta_i_logN_ordered$i)[order(theta_i_logN_ordered$.value)]
                     )
  ) |>
  rename(theta = .value) |>
  ggplot() +
  aes(x = ID, y = theta) +
  geom_hline(yintercept = 0, color = "black", linetype = "22") + #string-linetype: line for 2 units, followed by 2 off
  stat_pointinterval(           #draws point est + CI for draw-input, point_interval function is median_qi by default
    fill = viridisLite::inferno(1, begin = 0.9), interval_color = viridisLite::inferno(1, begin = 0.9), 
    point_color = "white", .width = c(.95), fatten_point = 2, linewidth = 1, shape = 21
  ) +
  labs(y = expression(theta[i]), x = "Participant ID") +
  coord_cartesian(ylim = c(0,7)) +
  papaja::theme_apa(base_size = 16) + 
  theme(
    axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 0)
  )
  
```

```{r}
#| label: fig-post-population-distribution-log-normal

post_populDist_logN <- tidybayes::gather_draws(
  fle_logN_bayes
  , mu_theta, sigma_theta
) |>  
  tidyr::nest(data = c(.chain, .iteration, .draw)) |>
  tidyr::pivot_wider(names_from = .variable, values_from = .value) |>
  mutate(
    density = purrr::map2(                      # iteration of each density function for each pair of mu and sigma
      mu_theta, sigma_theta, 
      \(mu, sigma){
        dlnorm(
          x = seq(0, 7, length.out = 101)
          , meanlog = mu
          , sdlog = sigma
        )
      }
    )
    ,theta = list(seq(0, 7, length.out = 101))    # corresponsing values on x for plot
  )
  
 theta_logN_dist <- post_populDist_logN |>
  tidyr::unnest(cols = c(density, theta)) |>
  ggplot() + 
  aes(x = theta, y = density) +
   geom_vline(xintercept = 0, color = "black", linetype = "22") +
   geom_hline(yintercept = 0, color = "black") + 
   geom_lineribbon(                              # function designed to use with output from point_interval
     aes(ymin = after_stat(ymin), ymax = after_stat(ymax))
     , stat = "summary"
     , fun.data = "median_qi"
     , color = "black"
     , fill = "black"
     , alpha = 0.2
   ) + 
   labs(x = expression(theta[i]), y = "Density") + 
   papaja::theme_apa(base_size = 16)

```

```{r}
#| label: fig-posterior-theta-distribution-log-normal
#| fig-cap: "Posterior distribution of individual effects $\\theta_i$ for each participant (left) and the posterior distribution of the population distribution of individual effects $\\theta_i$ (right). The horizontal dashed line marks the point of no effect, i.e., $\\theta_i = 0$."

log_normal_caterpillar + (theta_logN_dist + coord_flip() + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())) +
  plot_layout(ncol = 2, width = c(1, 1/3), axes = "collect_y")
```

### Bayes factor with unconstrained model

```{r}
#| label: bayes-factor-logN

logN_ml <- bridge_sampler(fle_logN_bayes, method = "warp3", repetitions = 30, silent = TRUE)
bayes_factor(unconstrained_ml, logN_ml)
bayes_factor(logN_ml, dominance_ml) -> bf_logN_dom
```

We still find decisive evidence in favor of an unconstrained distribution of $\theta_i$. As the log-normal model provides a better fit to the data compared to the model assuming a truncated normal distribution for $\theta_i$, the Bayes factor for the unconstrained model over the log-normal model is approximately 1/`r round(bf_logN_dom$bf_median_based,0)` of the Bayes factor for the unconstrained model over the truncated normal model.


## Gamma distributed $\theta_i$

Another potential distribution is the Gamma distribution.
As before, priors are placed on the mean ($\bar{x}$) and the standard deviation ($s$) of the Gamma distribution.

$$
\begin{aligned}
y_{ij} & \sim \mathcal{N}(\theta_i, \sigma_\epsilon) \\ \\
\theta_i & \sim \log \mathcal{G}(\alpha_\theta, \beta_\theta) \\
\sigma_\epsilon^2 & \propto 1/\sigma_\epsilon^2 \\ \\
\alpha_\theta & = \frac{\bar{x}^2}{s^2} \\
\beta_\theta & = \frac{\bar{x}}{s^2} \\ \\
\bar{x} & \sim \log \mathcal{N}(a_x, b_x) \\
s & \sim \log \mathcal{N}(a_s, b_s)
\end{aligned}
$$


```{r}
#| label: fit-gamma-model

pse_gamma_model <- readLines("./models/pse-dominance-gamma.stan") |>
  paste(collapse = "\n")

fle_stan_gamma_data <- list(
  N = nrow(fle_data) |> as.integer()
  , n_subjects = length(unique(fle_data$ID)) |> as.integer()
  , id = as.integer(as.factor(fle_data$ID))
  , y = fle_data$dva
  
  # Prior settings
  , mean_gamma_mu = 0.5
  , mean_gamma_sigma = 1
  , sd_gamma_mu = 0.5
  , sd_gamma_sigma = 1
)

fle_gamma_bayes <- rstan::stan(
  model_code = pse_gamma_model
  , data = fle_stan_gamma_data
  , iter = 1e5/4 + 1000
  , warmup = 1000
  , chains = 4
  , cores = 4
  , seed = 123
)

# model diagnostics
str_post_gamma <- data.frame(summary(fle_gamma_bayes)$summary)
check_hmc_diagnostics(fle_gamma_bayes)
```

### Estimated of individual effects 

```{r}
#| label: fig-indiv-effects-gamma

theta_i_gamma <- tidybayes::gather_draws(
  fle_gamma_bayes
  , theta[i]
)

theta_i_gamma_ordered <- tidybayes::median_qi(theta_i_gamma) |>
  mutate(i = factor(i
                    ,levels = unique(i)[order(.value)]        
                    ,labels = unique(i)[order(.value)]        #I create this solely for creating levels in draws, as geom-f calls on draws
                    ) 
         )

# Plot median estimates
gamma_caterpillar <- theta_i_gamma |>
  mutate(ID = factor(i
                     ,levels = unique(theta_i_gamma_ordered$i)[order(theta_i_gamma_ordered$.value)]  #new var ID with ordered med levels
                     ,labels = unique(theta_i_gamma_ordered$i)[order(theta_i_gamma_ordered$.value)]
                     )
  ) |>
  rename(theta = .value) |>
  ggplot() +
  aes(x = ID, y = theta) +
  geom_hline(yintercept = 0, color = "black", linetype = "22") + #string-linetype: line for 2 units, followed by 2 off
  stat_pointinterval(           #draws point est + CI for draw-input, point_interval function is median_qi by default
    fill = viridisLite::inferno(1, begin = 0.9), interval_color = viridisLite::inferno(1, begin = 0.9), 
    point_color = "white", .width = c(.95), fatten_point = 2, linewidth = 1, shape = 21
  ) +
  labs(y = expression(theta[i]), x = "Participant ID") +
  coord_cartesian(ylim = c(0,7)) +
  papaja::theme_apa(base_size = 16) + 
  theme(
    axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 0)
  )
  
```

```{r}
#| label: fig-post-population-distribution-gamma

post_populDist_gamma <- tidybayes::gather_draws(
  fle_gamma_bayes
  , alpha_theta, beta_theta
) |>  
  tidyr::nest(data = c(.chain, .iteration, .draw)) |>
  tidyr::pivot_wider(names_from = .variable, values_from = .value) |>
  mutate(
    density = purrr::map2(                      # iteration of each density function for each pair of mu and sigma
      alpha_theta, beta_theta, 
      \(alpha, beta){
        dgamma(
          x = seq(0, 7, length.out = 101)
          , shape = alpha
          , rate = beta
        )
      }
    )
    ,theta = list(seq(0, 7, length.out = 101))    # corresponsing values on x for plot
  )
  
 theta_gamma_dist <- post_populDist_gamma |>
  tidyr::unnest(cols = c(density, theta)) |>
  ggplot() + 
  aes(x = theta, y = density) +
   geom_vline(xintercept = 0, color = "black", linetype = "22") +
   geom_hline(yintercept = 0, color = "black") + 
   geom_lineribbon(                              # function designed to use with output from point_interval
     aes(ymin = after_stat(ymin), ymax = after_stat(ymax))
     , stat = "summary"
     , fun.data = "median_qi"
     , color = "black"
     , fill = "black"
     , alpha = 0.2
   ) + 
   labs(x = expression(theta[i]), y = "Density") + 
   papaja::theme_apa(base_size = 16)

```

```{r}
#| label: fig-posterior-theta-distribution-gamma
#| fig-cap: "Posterior distribution of individual effects $\\theta_i$ for each participant (left) and the posterior distribution of the population distribution of individual effects $\\theta_i$ (right). The horizontal dashed line marks the point of no effect, i.e., $\\theta_i = 0$."

gamma_caterpillar + (theta_gamma_dist + coord_flip() + theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())) +
  plot_layout(ncol = 2, width = c(1, 1/3), axes = "collect_y")
```

### Bayes factor with unconstrained model

```{r}
#| label: bayes-factor-gamma

gamma_ml <- bridge_sampler(fle_gamma_bayes, method = "warp3", repetitions = 30, silent = TRUE)
bayes_factor(unconstrained_ml, gamma_ml)
bayes_factor(gamma_ml, dominance_ml) -> bf_gamma_dom
```

We still find decisive evidence in favor of an unconstrained distribution of $\theta_i$. As the Gamma model provides a better fit to the data compared to the model assuming a truncated normal distribution for $\theta_i$, the Bayes factor for the unconstrained model over the Gamma model is approximately 1/`r round(bf_gamma_dom$bf_median_based,0)` of the Bayes factor for the unconstrained model over the truncated normal model.


## Spike-and-slab distributed $\theta_i$

Another potential distribution is the spike-and-slab distribution. The spike-and-slab distribution is a mixture of two distributions, one of which is a point mass at zero (the spike; no effect) and the other is a continuous distribution (the slab; flash lag or lead effect).
As before, priors are placed on the mean ($\bar{x}$) and the standard deviation ($s$) of the Gamma distribution.

```{r}
#| label: fit-spike-and-slab-model

pse_unconstrained_spike_model <- readLines("./models/pse-unconstrained-spike.stan") |>
  paste(collapse = "\n")

fle_stan_unconstrained_spike_data <- list(
  N = nrow(fle_data) |> as.integer()
  , n_subjects = length(unique(fle_data$ID)) |> as.integer()
  , id = as.integer(as.factor(fle_data$ID))
  , y = fle_data$dva
  
  # Prior settings
  , mu_theta_mean = 0.5
  , mu_theta_sd = 1
  , pi_mean = 0.5
  , pi_sd = 1
  , omega_mean = 0
  , omega_sd = 0.5
)

fle_unconstrained_spike_bayes <- rstan::stan(
  model_code = pse_unconstrained_spike_model
  , data = fle_stan_unconstrained_spike_data
  , iter = 1e5/4 + 1000
  , warmup = 1000
  , chains = 4
  , cores = 4
  , seed = 123
)

# model diagnostics
str_post_unconstrained_spike <- data.frame(summary(fle_unconstrained_spike_bayes)$summary)
check_hmc_diagnostics(fle_unconstrained_spike_bayes)
```

```{r}
#| label: spike-slab-posteriors

posterior_population_parameters <- tidybayes::gather_draws(
  fle_unconstrained_spike_bayes
  , mu_theta, sigma_theta, p_negative_theta, omega
) |>
  dplyr::mutate(
    .variable = factor(
      .variable
      , levels = c("mu_theta", "sigma_theta", "p_negative_theta", "omega")
      , labels = c("mu[theta]", "sigma[theta]", "p(theta[i] <= 0)", "p(theta[i] == 0 | p(theta[i] <= 0))")
    )
  )

(posterior_population_parameters |>
  ggplot() +
  aes(x = .value, fill = after_stat(level)) +
  stat_halfeye(fill_type = "segments", normalize = "panels", .width = c(.5, .8, .95, .99, 1), slab_color = "black", alpha = 0.7, fatten_point = 2, interval_size_range = c(1, 2)) +
  annotate("point", x = c(0, 0.5), y = 0, color = "white") + 
  scale_fill_viridis_d(option = "B", begin = 0.9, end = 0.1) +
  guides(
    fill = guide_legend(
      title = "Credible level"
      , override.aes = list(color = NA)
      , reverse = TRUE
      , direction = "horizontal"
      , title.position = "top"
      , title.hjust = 0.5
    )
  ) +
  coord_cartesian(ylim = c(-0.05, NA)) +
  facet_wrap(
    ~ .variable
    , scales = "free"
    , labeller = label_parsed
    , ncol = 2
  ) +
  labs(x = "Parameter value") +
  papaja::theme_apa(base_size = 16) +
  theme(
    axis.line.y = element_blank()
    , axis.ticks.y = element_blank()
    , axis.text.y = element_blank()
    , axis.title.y = element_blank()
    # , legend.title.position = "top"
  )) |>
  lemon::reposition_legend("center", panel = 'panel-2-2')
```

```{r}
#| label: joint-posterior-spike-slab

posterior_population_parameters |>
  dplyr::filter(.variable %in% c("p(theta[i] <= 0)", "p(theta[i] == 0 | p(theta[i] <= 0))")) |>
  tidyr::pivot_wider(names_from = .variable, values_from = .value) |>
  dplyr::mutate(
    `p(theta[i] == 0)` = `p(theta[i] <= 0)` * `p(theta[i] == 0 | p(theta[i] <= 0))`
  ) |>
  ggplot() +
    aes(x = `p(theta[i] <= 0)`, y = `p(theta[i] == 0 | p(theta[i] <= 0))`, color = `p(theta[i] == 0)`) +
    geom_point(alpha = 0.1) +
    stat_density_2d(color = "white", bins = 10) +
    geom_abline(intercept = 0, slope = 1, color = "white", linetype = "22") +
    # annotate("point", x = fixef(fle_lmer)[1], y = attr(VarCorr(fle_lmer)$ID, "stddev"), shape = 21, color = "black", fill = "white", size = 5) +
    labs(
      x = expression(p(theta[i] <= 0))
      , y = expression(p(theta[i] == 0*"|"*p(theta[i] <= 0)))
      , color = expression(p(theta[i] == 0))
    ) +
    scale_color_viridis_c(option = "C", limits = c(0, 0.5)) +
    lims(x = c(0, 0.5), y = c(0, 1)) +
    papaja::theme_apa(base_size = 16)
```

### Bayes factor with unconstrained model

```{r}
#| label: bayes-factor-spike-and-slab

unconstrained_spike_ml <- bridge_sampler(fle_unconstrained_spike_bayes, method = "warp3", repetitions = 30, silent = TRUE)
bayes_factor(unconstrained_ml, unconstrained_spike_ml)
bayes_factor(unconstrained_spike_ml, dominance_ml) -> bf_unconstrained_spike_dom
```

We still find decisive evidence in favor of an unconstrained distribution of $\theta_i$. As the spike-and-slab model provides a better fit to the data compared to the model assuming a truncated normal distribution for $\theta_i$, the Bayes factor for the unconstrained model over the spike-and-slab model is approximately 1/`r round(bf_unconstrained_spike_dom$bf_median_based,0)` of the Bayes factor for the unconstrained model over the truncated normal model.

# Prior settings
## Prior settings for mean of $\theta_i$

To assess the (possible) impact of different prior assumptions on the results of the model comparison, the prior distributions for the mean of $theta_i$ are plotted below. 
Since the common model assumes one fixed effect shared by all individuals, the prior on the common effect can be understood as a prior on the mean effect. 

```{r}
#| label: prior-mean-cross-model-comparison
#| fig-cap: "Comparison of priors for $\\mu_{\\theta}$ across the unconstrained, log-normal and common model (truncated tbd)"

# Calculate prior implied for mean of theta_i of truncated normal model
trunc_alpha <- (0-mu_prior_truncn)/sigma_prior_truncn
trunc_Z <- 1-pnorm(trunc_alpha)
mean_prior_truncn <- mu_prior_truncn + sigma_prior_truncn*(dnorm(trunc_alpha)/trunc_Z)

# Plot
data.frame(priors = c(mu_prior
                      ,mean_prior_log
                      ,mean_prior_truncn
                      ,mu_prior_common)
                      ,model = rep(c("unconstrained","log-normal","truncated normal","common"), each = 1e5)) |>
  ggplot() +
  aes(x = priors, fill = model) +
  lims(x = c(0,5)) +
  geom_density(alpha = 0.2, linewidth = 0.5) +
  labs(
    x = expression(mu[theta])
    , y = "Density"
    , fill = "Prior for"
  ) +
  scale_color_viridis_d(option = "B", begin = 0.1, end = 0.8, aesthetics = "fill") +
  papaja::theme_apa(base_size = 16)
   
```

## Prior settings for standard deviation of $\theta_i$

To assess the (possible) impact of different prior assumptions on the results of the model comparison, the prior distributions for the standard deviation of $theta_i$ are plotted below. 


```{r}
#| label: prior-sd-cross-model-comparison
#| fig-cap: "Comparison of priors for $\\sigma_{\\theta}$ across the unconstrained, log-normal and average null model (truncated tbd)"

# Calculate implied prior for sd of theta_i of truncated normal model
sd_prior_truncn <- sqrt(
  (sigma_prior_truncn^2)*
    (1 + trunc_alpha*dnorm(trunc_alpha)/trunc_Z -(dnorm(trunc_alpha)/trunc_Z)^2)
)

# Plot
data.frame(priors = c(sigma_prior
                      ,sd_prior_log
                      ,sd_prior_truncn
                      ,sigma_prior_average)
                      ,model = rep(c("unconstrained","log-normal","truncated normal","average null"), each = 1e5)) |>
  ggplot() +
  aes(x = priors, fill = model) +
  lims(x = c(0,5)) +
  geom_density(alpha = 0.2, linewidth = 0.5) +
  labs(
    x = expression(sigma[theta])
    , y = "Density"
    , fill = "Prior for"
  ) +
  scale_color_viridis_d(option = "G", begin = 0.1, end = 0.8, aesthetics = "fill") +
  papaja::theme_apa(base_size = 16)
   
```

# Bayesian model comparison

## Bayes factor summary table

```{r}
#| label: bf-summary-table
#| tbl-cap: "Summary table: Bayes Factors over unconstrained model (=reference model) and naive posterior probability (NPP), based on assumption of equal prior model odds. Maximum $\\hat{R}$ and coefficient of variation (CV) for the bridge sampling are also given for each model."

bfs_base_unc <- c(bayes_factor(null_ml, unconstrained_ml)$bf_median_based
    , bayes_factor(average_ml, unconstrained_ml)$bf_median_based
    , bayes_factor(common_ml, unconstrained_ml)$bf_median_based
    , bayes_factor(unconstrained_spike_ml, unconstrained_ml)$bf_median_based
    , bayes_factor(logN_ml, unconstrained_ml)$bf_median_based
    , bayes_factor(gamma_ml, unconstrained_ml)$bf_median_based
    , bayes_factor(dominance_ml, unconstrained_ml)$bf_median_based
    , 1
    )

# Naive posterior probability: naive assumption that prior odds = 1 for all model comparisons
# -> post prob = bf / sum(bf)

npp_base_unc <- bfs_base_unc / sum(bfs_base_unc) 

# model diagnostics
rhat <- c(max(str_post_null$Rhat)
          ,max(str_post_average$Rhat)
          ,max(str_post_common$Rhat)
          ,max(str_post_unconstrained_spike$Rhat)
          ,max(str_post_logN$Rhat)
          ,max(str_post_gamma$Rhat)
          ,max(str_post_dominance$Rhat)
          ,max(str_post_unconstrained$Rhat))


# bridge sampling CV
 bridge.cv <- c(sd(null_ml$logml)/abs(mean(null_ml$logml))
                ,sd(average_ml$logml)/abs(mean(average_ml$logml))
                ,sd(common_ml$logml)/abs(mean(common_ml$logml))
                ,sd(unconstrained_spike_ml$logml)/abs(mean(unconstrained_spike_ml$logml))
                ,sd(logN_ml$logml)/abs(mean(logN_ml$logml))
                ,sd(gamma_ml$logml)/abs(mean(gamma_ml$logml))
                ,sd(dominance_ml$logml)/abs(mean(dominance_ml$logml))
                ,sd(unconstrained_ml$logml)/abs(mean(unconstrained_ml$logml)))

# summary table
data.frame(
  model = c("Null model", "Average null model", "Common model", "Spike-and-slab model", "Log-normal model", "Gamma model", "Truncated normal model", "Unconstrained model")
  , bf = printnum(bfs_base_unc, format = "e")
  , npp = printnum(npp_base_unc, gt1 = F, digits = 2)
  , rhat = printnum(rhat, digits = 3)
  , cv = printnum(bridge.cv, format = "e")
)  |> knitr::kable(
  col.names = c("Model ($i$)", "$BF_{i/unconstrained}$", "NPP", "max.$\\hat{R}$", "Bridge Sampling CV")
)
```

In summary, we find overwhelming evidence for the unconstrained model and the existence of not only flash lag, but also flash lead effects.
Given the data at hand, the posterior probability that no one exhibits a flash lead effect is practically zero (with a priori assumption that all models are equally likely).

```{r}
#| label: bf-model-averaging

bf_dom_null <- c(bayes_factor(logN_ml,null_ml)$bf_median_based
                 , bayes_factor(gamma_ml,null_ml)$bf_median_based
                 , bayes_factor(dominance_ml, null_ml)$bf_median_based)
bf_unc_null <- bayes_factor(unconstrained_ml, null_ml)$bf_median_based

av_bf_dom_unc <- length(bf_dom_null)^(-1)*(sum(bf_dom_null)) / bf_unc_null

```

To summarize the results above, we used Bayesian model averaging to compare the unconstrained model with all dominance models collectively.
Assuming the unconstrained model and the set of dominance models are equally likely a priori, $BF_{dom/unc} =$ `r printnum(av_bf_dom_unc, format = "e")`. \
Thus, the data support the unconstrained model over the set of dominance models by a factor of `r printnum(av_bf_dom_unc^(-1),format = "f")`-to-1. 
